---
title: "Sampling Chap 5"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

5. In Chapter 4, we used logistic regression to predict the probability of default using income and balance on the Default data set. We will now estimate the test error of this logistic regression model using the validation set approach. Do not forget to set a random seed before beginning your analysis.

  a - Fit a logistic regression model that uses income and balance to
predict default.

```{r}
set.seed(123)
library(ISLR)
data(Default)
glm_default <- glm(default ~  income + balance, data = Default, family = binomial)
summary(glm_default)
```


  b - Using the validation set approach, estimate the test error of this model. In order to do this, you must perform the following steps:
    i. Split the sample set into a training set and a validation set.

``` {r}
set.seed(123)
trainIndex <- sample(nrow(Default), nrow(Default) * .5)

#train <- Default[trainIndex, ]
#test <- Default[-trainIndex, ]
```

  ii. Fit a multiple logistic regression model using only the training observations.

```{r}
glm_def <- glm(default ~ income + balance, data = Default, family = binomial, subset = trainIndex)
summary(glm_def)
```

  iii. Obtain a prediction of default status for each individual in the validation set by computing the posterior probability of default for that individual, and classifying the individual to the default category if the posterior probability is greater than 0.5.

```{r}
default_pred <- predict(glm_def, newdata = Default[-trainIndex, ], type = 'response')
default_prob <- ifelse(default_pred > .5, 'Yes', 'No')
table(Default[-trainIndex, ]$default, default_prob)
```

  iv. Compute the validation set error, which is the fraction of the observations in the validation set that are misclassified.
```{r}
mean(Default[-trainIndex, ]$default != default_prob)
```

  c - Repeat the process in (b) three times, using three different splits of the observations into a training set and a validation set. Comment on the results obtained.

```{r}
set.seed(123)
trainIndex <- sample(nrow(Default), nrow(Default) * .5)
glm_def <- glm(default ~ income + balance, data = Default, family = binomial, subset = trainIndex)
default_pred <- predict(glm_def, newdata = Default[-trainIndex, ], type = 'response')
default_prob <- ifelse(default_pred > .5, 'Yes', 'No')
table(Default[-trainIndex, ]$default, default_prob)
mean(Default[-trainIndex, ]$default != default_prob)


set.seed(456)
trainIndex <- sample(nrow(Default), nrow(Default) * .5)
glm_def <- glm(default ~ income + balance, data = Default, family = binomial, subset = trainIndex)
default_pred <- predict(glm_def, newdata = Default[-trainIndex, ], type = 'response')
default_prob <- ifelse(default_pred > .5, 'Yes', 'No')
table(Default[-trainIndex, ]$default, default_prob)
mean(Default[-trainIndex, ]$default != default_prob)

set.seed(789)
trainIndex <- sample(nrow(Default), nrow(Default) * .5)
glm_def <- glm(default ~ income + balance, data = Default, family = binomial, subset = trainIndex)
default_pred <- predict(glm_def, newdata = Default[-trainIndex, ], type = 'response')
default_prob <- ifelse(default_pred > .5, 'Yes', 'No')
table(Default[-trainIndex, ]$default, default_prob)
mean(Default[-trainIndex, ]$default != default_prob)
```

  - Test error seems to be consistent ~ 2.7%.
  
  d - Now consider a logistic regression model that predicts the probability of default using income, balance, and a dummy variable for student. Estimate the test error for this model using the validation set approach. Comment on whether or not including a
dummy variable for student leads to a reduction in the test error rate

```{r}
set.seed(123)
trainIndex <- sample(nrow(Default), nrow(Default) * .5)
glm_def <- glm(default ~ ., data = Default, family = binomial, subset = trainIndex)
default_pred <- predict(glm_def, newdata = Default[-trainIndex, ], type = 'response')
default_prob <- ifelse(default_pred > .5, 'Yes', 'No')
table(Default[-trainIndex, ]$default, default_prob)
mean(Default[-trainIndex, ]$default != default_prob)
```

- Both models give roughly the same result.

6. We continue to consider the use of a logistic regression model to predict the probability of default using income and balance on the Default data set. In particular, we will now compute estimates for the standard errors of the income and balance logistic regression coefficients in two different ways: 
  (1) using the bootstrap, and 
  (2) using the standard formula for computing the standard errors in the glm()
function. 

    a - Using the summary() and glm() functions, determine the estimated standard errors for the coefficients associated with income and balance in a multiple logistic regression model that uses both predictors.
    
```{r}
set.seed(123)
default_glm1 <- glm(default ~ income + balance, data = Default, family = binomial)
summary(default_glm1)$coef
```
    - SE for income = 4.985167e-06
    - SE for balance = 2.273731e-04
    
    b - Write a function, boot.fn(), that takes as input the Default data set as well as an index of the observations, and that outputs the coefficient estimates for income and balance in the multiple logistic regression model.

```{r}
set.seed(123)
boot.fn <- function(Default, trainIndex) {
  return(glm(default ~ income + balance, data = Default, family = binomial, subset = trainIndex)$coef)
}
```

c - Use the boot() function together with your boot.fn() function to estimate the standard errors of the logistic regression coefficients for income and balance.

```{r}
boot.fn(Default, 1:nrow(Default))
library(boot)
boot(Default, boot.fn, R = 100)
```

d - Comment on the estimated standard errors obtained using the glm() function and using your bootstrap function.

  - SE for income = 2.080898e-05 using glm and = 2.080898e-05
  - SE for balance = 5.647103e-03 using glm and = 5.647103e-03
  
  The results are pretty close using both methods.
  

7. In Sections 5.3.2 and 5.3.3, we saw that the cv.glm() function can be used in order to compute the LOOCV test error estimate. Alternatively, one could compute those quantities using just the glm() and predict.glm() functions, and a for loop. You will now take this approach in order to compute the LOOCV error for a simple logistic regression model on the Weekly data set. Recall that in the context of classification problems, the LOOCV error is given in (5.4).

(a) Fit a logistic regression model that predicts Direction using Lag1 and Lag2.

```{r}
glm_weekly <- glm(Direction ~ Lag1 + Lag2, data = Weekly, family = binomial)
```

(b) Fit a logistic regression model that predicts Direction using Lag1 and Lag2 using all but the first observation.

```{r}
set.seed(123)
glm_weekly <- glm(Direction ~ Lag1 + Lag2, data = Weekly[-1, ], family = binomial)
```

(c) Use the model from (b) to predict the direction of the first observation. You can do this by predicting that the first observation will go up if P(Direction="Up"|Lag1, Lag2) > 0.5. Was this observation correctly classified?

```{r}
set.seed(123)
glm_pred <- ifelse(predict(glm_weekly, newdata = Weekly[1, ], type = 'response') > .5, 'Up', 'Down')
```

(d) Write a for loop from i = 1 to i = n, where n is the number of observations in the data set, that performs each of the following steps:

  i. Fit a logistic regression model using all but the ith observation to predict Direction using Lag1 and Lag2.
  ii. Compute the posterior probability of the market moving up for the ith observation.
  iii. Use the posterior probability for the ith observation in order to predict whether or not the market moves up.
  iv. Determine whether or not an error was made in predicting the direction for the ith observation. If an error was made, then indicate this as a 1, and otherwise indicate it as a 0.

```{r}
set.seed(123)
loocv_err <- rep(0, nrow(Weekly))
for (i in 1:nrow(Weekly)) {
  # Fit glm model to the Weekly data
  glm_weekly <- glm(Direction ~ Lag1 + Lag2, data = Weekly[-i, ], family = binomial)
  
  # Predit the unseen obs i
  glm_pred <- ifelse(predict(glm_weekly, newdata = Weekly[1, ], type = 'response') > .5, 'Up', 'Down')
  
  loocv_err[i] <- ifelse(Weekly[i, ]$Direction == glm_pred, 0, 1)
}

str(loocv_err)
```

  e - Take the average of the n numbers obtained in (d)iv in order to obtain the LOOCV estimate for the test error. Comment on the results.

```{r}
mean(loocv_err)
```

  - The estimated test error = 44.4%
  

8. We will now perform cross-validation on a simulated data set.
  a - Generate a simulated data set as follows:
  
```{r}
set.seed(1)
x <-  rnorm(100)
y <- x - 2 * x ^ 2 + rnorm(100)
```
  
In this data set, what is n and what is p? Write out the model used to generate the data in equation form.

  => `n` = 100
  => `p` = 2 variables

The model used to generate the data is : `y = x - 2 * x ^ 2 + epsilon`.

  b - Create a scatterplot of X against Y . Comment on what you find.
  
```{r}
plot(x, y)
```

  => The relationship between x and y is rather quadratic.

  c - Set a random seed, and then compute the LOOCV errors that result from fitting the following four models using least squares:
    i. Y = ??0 + ??1X + epsilon
    ii. Y = ??0 + ??1X + ??2X2 + epsilon
    iii. Y = ??0 + ??1X + ??2X2 + ??3X3 + epsilon
    iv. Y = ??0 + ??1X + ??2X2 + ??3X3 + ??4X4 + epsilon

  
```{r}
set.seed(123)
df <- data.frame(y, x, x2 = x ^ 2, x3 = x ^ 3, x4 = x ^ 4)

df_coef <- 
cv.err <- rep(0, 4)
for (i in 1:4) {
  glm.fit <- glm(y ~ poly(x, i), data = df)
  cv.err[i] <- cv.glm(df, glm.fit)$delta[1]
}

cv.err
```

  d - Repeat (c) using another random seed, and report your results. Are your results the same as what you got in (c)? Why?

```{r}
set.seed(456)

df <- data.frame(y, x, x2 = x ^ 2, x3 = x ^ 3, x4 = x ^ 4)

cv.err2 <- rep(0, 4)
for (i in 1:4) {
  glm.fit <- glm(y ~ poly(x, i), data = df)
  cv.err2[i] <- cv.glm(df, glm.fit)$delta[1]
}

cv.err2
```

  => The results are exactly the same because LOOCV predicts every observation ,,,,,,,,,,,,
  
  e - Which of the models in (c) had the smallest LOOCV error? Is this what you expected? Explain your answer.
  
  => The quadratic model has the lowest LOOCV error. This is expected as a quadratic formula is used to generate the data.

  f  - Comment on the statistical significance of the coefficient estimates that results from fitting each of the models in (c) using least squares. Do these results agree with the conclusions drawn based on the cross-validation results?

```{r}
summary(glm(y ~ poly(x, 4), data = df))
```

  - `x and x ^ 2 are statistically significant`, however, x ^ 3 and x ^ 4 are not. This confirms LOOCV result.


  9 - We will now consider the Boston housing data set, from the MASS library.

  a - Based on this data set, provide an estimate for the population mean of medv. Call this estimate ????.
  
```{r}
library(MASS)
u_hat <- mean(Boston$medv)
```

  b - Provide an estimate of the standard error of ????. Interpret this result.
  Hint: We can compute the standard error of the sample mean by dividing the sample standard deviation by the square root of the number of observations.

```{r}
sd(Boston$medv) / sqrt(nrow(Boston))
```

  c - Now estimate the standard error of ???? using the bootstrap. How does this compare to your answer from (b)?

```{r}
set.seed(123)

mean_fc <- function(col, i){
  return(mean(col[i]))
}

bootstrap_res <- boot(Boston$medv, mean_fc, R = 100)
bootstrap_res
```

  - `Both SE are relatively close (.408 vs .438).`
  
  d - Based on your bootstrap estimate from (c), provide a 95 % confidence interval for the mean of medv. Compare it to the results obtained using t.test(Boston$medv).
  Hint: You can approximate a 95 % confidence interval using the formula [???? ??? 2SE(????), ???? + 2SE(????)].

```{r}
confint_lower <- bootstrap_res$t0 - 2 * sd(bootstrap_res$t)
confint_upper <- bootstrap_res$t0 + 2 * sd(bootstrap_res$t)

confint_lower
confint_upper
t.test(Boston$medv)
```

  e - Based on this data set, provide an estimate, ????med, for the median value of medv in the population.
  
```{r}
median(Boston$medv)
```
  
  f - We now would like to estimate the standard error of ????med. Unfortunately, there is no simple formula for computing the standard error of the median. Instead, estimate the standard error of the median using the bootstrap. Comment on your findings.

```{r}
set.seed(123)

median_fc <- function(col, i){
  return(median(col[i]))
}

bootstrap_med <- boot(Boston$medv, median_fc, R = 100)
bootstrap_med
```

  => `Estimated SE for the median = .429`
  
  g - Based on this data set, provide an estimate for the tenth percentile of medv in Boston suburbs. 
  
```{r}
quantile(Boston$medv, probs = .1)
```

  h - Use the bootstrap to estimate the standard error of ????0.1. Comment on your findings.
  
```{r}
set.seed(123)

quant_fc <- function(col, i){
  return(quantile(col[i], probs = .1))
}

bootstrap_quant <- boot(Boston$medv, quant_fc, R = 100)
bootstrap_quant
```

=> `Estimated SE for the median = .539`



